---
title: "Land Use and Watersheds of the Big Island Hawai'i"
author: "Sandra Fogg"
date: "2/17/2020"
output: html_document
---

```{r setup, warning = FALSE, message = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE, include=FALSE}
library(tidyverse)
library(sf)
library(sp)
library(rgdal)
library(raster)
library(dplyr)
library(rgeos)
library(scales)
library(fasterize)
library(dismo)
library(maptools)
library(sdmpredictors)
library(leaflet)
library(maps)
library(ncdf4)
library(marmap)
library(rnaturalearth)
library(dplyr)
library(ggmap)
library(rmapshaper)
library(ggspatial)
library(paletteer)
```
## LANDCOVER
<br>
The Big Island of Hawai'i is one of the most biodiverse places on Earth. There are many different types of biomes and microclimates on the island that support an abundance of wild flora and fauna, as well as human uses. In this analysis, land use and cover is generalized and visualized into three separate categories: Agriculture, Urban/Residential/Commercial, and Natural. These categories were developed by combining 27 specified land uses and coverage types.
<br>
<br>

### Data Processing
```{r, warning = FALSE, message = FALSE}
#Read in and Explore Shapefiles
# Read in Land Use and Land Cover
# Plot, Check Coordinate Reference System, and Explore Types of Cover
lulc <- read_sf("Land_Use_Land_Cover_LULC.shp")
#plot(lulc)
#st_crs(lulc)
#unique(lulc$landcover)

# Land Cover Types
# [1] "Cropland and Pasture"                                                     
# [2] "Commercial and Services"                                                  
# [3] "Residential"                                                              
# [4] "Evergreen Forest Land"                                                    
# [5] "Other Urban or Built-up Land"                                             
# [6] "Mixed Rangeland"                                                          
# [7] "Industrial"                                                               
# [8] "Streams and Canals"                                                       
# [9] "Orchards, Groves, Vineyards, Nurseries and Ornamental Horticultural Areas"
# [10] "Shrub and Brush Rangeland"                                                
# [11] "Forested Wetland"                                                         
# [12] "Reservoirs"                                                               
# [13] "Nonforested Wetland"                                                      
# [14] "Bare Exposed Rock"                                                        
# [15] "Sandy Areas Other than Beaches"                                           
# [16] "Transportation, Communications and Utilities"                             
# [17] "Herbaceous Rangeland"                                                     
# [18] "Beaches"                                                                  
# [19] "Other Agricultural Land"                                                  
# [20] "Lakes"                                                                    
# [21] "Strip Mines, Quarries, and Gravel Pits"                                   
# [22] "Mixed Barren Land"                                                        
# [23] "Bays and Estuaries"                                                       
# [24] "Mixed Urban or Built-up Land"                                             
# [25] "Transitional Areas"                                                       
# [26] "0"                                                                        
# [27] "Industrial and Commercial Complexes"                                      
# [28] "Confined Feeding Operations"                                              

#Read in Watersheds
# Plot, Check Coordinate Reference System, and Explore Types of Cover
watersheds <- read_sf("Watersheds.shp")
#plot(watersheds)
#st_crs(watersheds)
#unique(watersheds)

#  Watershed Measurements
# [1] objectid                                                    
# [2] wuname                                                
# [3] wuc                                                             
# [4] hucarea                                                   
# [5] huc                                            
# [6] swma                                                          
# [7] area_sqmi                                                              
# [8] st_areasha                                                     
# [9] st_perimet

```


<br>
Combine Landcover Types
```{r, warning = FALSE, message = FALSE}
# Categorize into Agriculture, Urban/Residential/Commercial, Natural
lulc_3cats <- lulc %>% 
  mutate(
    landcover = case_when(
       landcover %in% c("Cropland and Pasture","Mixed Rangeland", "Orchards, Groves, Vineyards, Nurseries and Ornamental Horticultural Areas",
                       "Shrub and Brush Rangeland", "Herbaceous Rangeland", 
                       "Other Agricultural Land","Confined Feeding Operations") ~ "Agriculture",
      landcover %in% c("Commercial and Services", "Residential", "Other Urban or Built-up Land", "Industrial", "Reservoirs",
                       "Transportation, Communications and Utilities", "Strip Mines, Quarries, and Gravel Pits", "Mixed Urban or Built-up Land", 
                       "Transitional Areas", "Industrial and Commercial Complexes") ~ "Urban/Residential/Commercial",
      landcover %in% c("Evergreen Forest Land", "Streams and Canals", "Forested Wetland", "Nonforested Wetland", "Bare Exposed Rock",
                       "Sandy Areas Other than Beaches", "Beaches", "Lakes", "Mixed Barren Land", "Bays and Estuaries") ~ "Natural",
    TRUE ~ landcover)) %>% 
      filter(landcover != "0")

#unique(lulc_3cats$landcover)
#lulc_3cats
```
<br>

### Results: Landcover Map

```{r, warning = FALSE, message = FALSE}
#Create a Map of Landcover and Uses in Hawaii
#Create a Static Map of Generalized Landcover Types
lulc_plt <- ggplot(lulc_3cats) +
  geom_sf(data = lulc_3cats,  aes(fill = landcover), color = NA) +
  coord_sf(xlim = c(-156.5, -154.5), ylim = c(18.5, 20.5), expand = FALSE) +
  labs(fill = "Landcover Type", x = "Longtitude", y = "Latitude", title = "Land Use on the Big Island, Hawai'i") +
  scale_fill_paletteer_d("awtools::spalette") +
  theme_classic() +
  theme(panel.background = element_rect(fill = "lightskyblue2")) 
 
lulc_plt

```
<br>

## WATERSHEDS

<br>
Hawaii's unique Pacific Island climate brings lots of annual precipitation to the region. The unique volcanic mountanious terrain separates the flows of these waters to many different parts of the island in specified watersheds. The map below depicts watersheds and their respective areas in square mileage.
<br>
<br>

### Results: Watersheds Map

```{r, warning = FALSE, message = FALSE}
#Create a Map of Watershed Areas
#Create a Static Map of Generalized Landcover Types
watersheds_plt <- ggplot(watersheds) +
  geom_sf(data = watersheds, aes(fill = watersheds$area_sqmi), color = "black") +
  coord_sf(xlim = c(-156.5, -154.5), ylim = c(18.5, 20.5), expand = FALSE) +
  labs(fill = "Watershed Size (Sq. Miles)", x = "Longtitude", y = "Latitude", title = "Watersheds on the Big Island Hawai'i") +
  scale_fill_paletteer_c("pals::ocean.matter") +
  theme_classic() +
  theme(panel.background = element_rect(fill = "lightskyblue")) 


watersheds_plt
```

